\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-21 concepts.tex
updates to intro etc

\subsubsection{\eu Plan Representation}
\label{sec:europa:pr}

@@

  \eu's ability to represent and reason over numeric intervals is a
  very useful characteristic that allow users to create flexible
  plans.

@@

\item \textit{Timelines}: Often objects in a domain must be described
  by exactly one token for every given timepoint in the plan. Any
  instances of a class derived from \eu's built-in Timeline class will
  induce ordering requirements among its tokens to ensure no temporal
  overlap may occur among them \cite{mus94}.  \comment{Maybe: add
    small example?}

@@

\item[\textbf{Token State Model}] In \eu's representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  Tokens.  Goals can be posted explicitly, for instance, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a result of domain
  compatibilities (described in the CAIP framework above), for
  instance, when the Shopping Agent must get to a Location where a
  Product is available before it is able to purchase it, in this case
  the original goal of purchasing the product will spawn a subgoal of
  being at a specific location.

@@

If neither activating nor merging works, a token may become REJECTED,
this is possible only for goals posted by the end user and specified
as optional. Subgoals generated as a result of token activation cannot
be rejected, if a subgoal cannot be satisfied through activation or
merging then it will cause the original activation decision that
spawned it to be retracted (\eu's search mechanisms are explained in
detail in a separate section later on).

>>>>

\section{\eu Basics}
\label{sec:basics}

\paragraph {A brief note on \eu history and notation} The earliest
origin of what we will call \eu in this chapter, is from the
\texttt{HSTS} planner \cite{mus94} which was used for a number of
demonstrations for space telescopes as noted in Section
\ref{sec:rabeyond}. \texttt{HSTS} was also the
\texttt{LISP}-based\footnote{This was the first (and only to our
  knowledge) software ever to been written in \texttt{LISP} to be
  flown in space.} planner that was flown as part of the Remote Agent
\texttt{RAX} experiment. Concurrent to \texttt{RAX}, a C++ version of
the planner was implemented at which went on to be used for the
\texttt{MAPGEN} \cite{bresina05} system and continues to be used to
date, in the mission-critical uplink process for the command/control
of the Mars Exploration Rovers (\texttt{MER}) mission. A vastly
re-factored, higher performance open source version \cite{europapso}
was named \eut and is the subject of this chapter and used onboard
MBARI's Dorado platform. Note however, in using the name \eu, we refer
to the \emph{fundamental ideas as well as the planning paradigm} even
if there exists a specific implementation.

\subsection{Introduction}
\label{sec:euintro}

\eu uses a \emph{domain model} written in a declarative language,
together with initial conditions and goals also in that specified
language, to construct a set of temporal relations that must be true
at the start time. These models include assertions about the physics
of the vehicle, i.e how it responds to external stimulus and
internally driven goals. By propagating these relations forward using
Simple Temporal Networks \cite{dechter91} and applying goal
constraints, \eu can select a set of conditions that should be true in
the future, where some of these conditions will correspond to actions
the agent must take. The planner can backtrack and try another path
during search if a goal cannot be reached while being capable of
discarding unachievable goals.

Traditionally robot execution has relied on dispatching commands at
precise times. Such linear sequences of precisely timed commands give
no ability to adjust execution on the basis of sensory
information. Although some commands can issue tests on sensor
readings, these tests have the objective of verifying whether expected
execution conditions are occuring. If not, the state of the system is
declared off-nominal and execution of the sequence of commands is
interrupted. More recently, executives have been proposed and
implemented that significantly broaden the way robots can be commanded
\cite{mus98,alami:1998p820}. For example, the Remote Agent executive
interpreted a \textit{temporally flexible plan} which represents each
start time as a variable and contains an explicit network of bounded
delay constraints between such variables.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\vskip+0.1cm
\end{figure}

Unlike a traditional fixed time-tagged command sequences, such
flexible plans leave room for adaptation at execution time. When the
executive considers when to start a task, it propagates information
through the constraint network, computes a time bound for the
variable, selects an actual execution time within the bound, and
starts the task at that time. Temporally flexible plans therefore,
express a \textit{range of possible outcomes} of the robots
interaction with the environment within which the executive can elect
at run time the most appropriate one for the actual execution
conditions. The fact that constraints are explicitly represented
ensures that through constraint propagation the executive will respect
global limits expressed in the plan (e.g., don't start a task until a
certain condition has been satisfied but still satisfying some global
deadline). Such flexibility is critical when dealing with dynamic
ocean conditions where precise timing of a robotic action might be
indeterminate. Further, the advantage of flexibility can be contrasted
with the consequences of the intrinsic inflexibiliy of traditional
command sequences. Because they are inflexible, sequences must
necessarily be designed considering worst case scenarios.

\eu uses a \emph{state variable} representation to describe the
evolution of state over time. The instantiated history of such state
variable evolution over a temporal horizon we call \emph{timelines}
and which represent a single thread in the execution of a concurrent
system. At any given time each thread can execute a single procedure.
Thus each timeline consists of a sequence of procedures which
encapsulate and describe state evolution; we call these instantiated
atomic entities \emph{tokens}.  A token therefore describes a
procedure invocation, the state variables on which it can occur, the
parameter values of the procedure, and the time values defining the
interval. We allow encapsulation of uncertainty within these tokens
with a range of start and end times and parameters, all of which are
encoded as variables. A constraint solver in turn manipulates these
variables defined in a \eu domain model. For example, consider
a scientific need to take a water samples 100 meters from a hotspot
while an AUV is performing a Yo-Yo in the water-column. Two samples
are needed if the feature's signal is above a threshold; one
otherwise. The token that is capturing the sensory threshold has a
parametric constraints to the token which fires the requisite water
sampler. In addition, the start time of the water sampling procedure
is highly dependant on the variability of sub-sea currents and actual
speed of the vehicle. Therefore a number of values are possible for
the start times and duration of the sampling all of which are valid
combinations for desired outcomes.

\subsection{\eu Plan Representation}
\label{sec:europa:pr}

@@

  \eus ability to represent and reason over numeric intervals is a
  very useful characteristic that allow users to create flexible
  plans.

@@

\item \textit{Timelines}: Often objects in a domain must be described
  by exactly one token for every given timepoint in the plan. Any
  instances of a class derived from \eus built-in Timeline class will
  induce ordering requirements among its tokens to ensure no temporal
  overlap may occur among them \cite{mus94}.  \comment{Maybe: add
    small example?}

@@

\item[\textbf{Token State Model}] In \eus representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  Tokens.  Goals can be posted explicitly, for instance, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a result of domain
  compatibilities (described in the CAIP framework above), for
  instance, when the Shopping Agent must get to a Location where a
  Product is available before it is able to purchase it, in this case
  the original goal of purchasing the product will spawn a subgoal of
  being at a specific location.

@@

If neither activating nor merging works, a token may become REJECTED,
this is possible only for goals posted by the end user and specified
as optional. Subgoals generated as a result of token activation cannot
be rejected, if a subgoal cannot be satisfied through activation or
merging then it will cause the original activation decision that
spawned it to be retracted (\eus search mechanisms are explained in
detail in a separate section later on).

### 2012-02-21 intro.tex
updates to intro etc

Persistent and early legitimate criticism of deliberation on the other
hand also well articulated by Brooks
\cite{Brooks91intelligencewithoutrea,Brooks91intelligencewithoutrep}
was targeting the performance of in-situ deliberation. Terrestrial
robotic platforms were well-known to frequently stop and ``plan''. In
the marine robotics world, such a situation would simply not be
permissible given the need to make continuous observations without
stopping an under-articulated robotic platform such as an AUV. However
algorithmic and implementational advances by the AI community over the
years has produced a range of planners with demonstrated embedded
capabilities \comment{cite}. Our work has been influenced by and in
turn influenced a wide range of activities in the field of AI Planning
and Plan Execution with demonstrated real-world and highly visible
capabilities in the Space domain
\cite{mus98,rajan00,aichang04,bresina05} for NASA. Lessons learned
have been applied to the world of marine robotics which we bring
together in this chapter. Important lessons that were derived from
this rich legacy involved the following:

@@

This chapter is organized as follows. We start with some foundational
concepts in our Planning framework in Section
\ref{sec:concepts}. \comment{add others} We conclude with Section
\ref{sec:conclusion}.

>>>>

Persistent and early legitimate criticism of deliberation on the other
hand also well articulated by Brooks
\cite{Brooks91intelligencewithoutrea,Brooks91intelligencewithoutrep}
was targeting the performance of in-situ deliberation. Terrestrial
robotic platforms were well-known to frequently stop and ``plan''. In
the marine robotics world, such a situation would simply not be
permissible given the need to make continuous observations without
stopping an under-articulated robotic platform such as an AUV. However
algorithmic and implementational advances by the AI community over the
years has produced a range of planners with demonstrated embedded
capabilities
\cite{simmons94,Haigh98,alami:1998p820,chien00,mus98,teichteil07}. Our
work has been influenced by and in turn influenced a wide range of
activities in the field of AI Planning and Plan Execution with
demonstrated real-world and highly visible capabilities in the Space
domain \cite{mus98,rajan00,aichang04,bresina05} for NASA. Lessons
learned have been applied to the world of marine robotics which we
bring together in this chapter. Important lessons that were derived
from this rich legacy involved the following:

@@

This chapter is organized as follows. We start with some foundational
concepts in our Planning framework in Section~\ref{sec:concepts} which
includes a brief overview of AI Planning in
Section~\ref{sec:planningfound}. In Section~\ref{sec:basics} we go
deeper into the \eu framework itself, the basis of all deliberation
within \rx. A key portion is Section~\ref{sec:europa:pr} which details
the Internal plan representation essential to planning within a
situated agent like \rx. Modeling (Section~\ref{sec:europa:modeling}),
Inference (Section~\ref{sec:europa:inference}) and Search
(Section~\ref{sec:europa:search}) further detail the important details
associated with \eu for the reader. With these foundational elements
out of the way, we transition to \rx in Section~\ref{sec:arch}, bring
out architectural details in Section~\ref{sec:arch:trex}, deal with
the execution cycle in Section~\ref{sec:arch:exec}, highlight how \rx
deliberates in Section~\ref{sec:arch:europa} finally delving deeper
into the key concept of synchronization in
Section~\ref{sec:arch:synch}. Further details on deliberation within
\rx (Section~\ref{sec:arch:plan}) and how planning and execution are
interleaved are shown in Section~\ref{sec:arch:intertwine}. We offer
results from our experiments at sea in Section~\ref{sec:results},
offer some insights into the near-term future goals of our research in
Section~\ref{sec:future} and conclude with
Section~\ref{sec:conclusion}.
